---
title: "Introduction to Statistical Concepts, MSCA 31000, Assignment 1b"
author: "Hope Foster-Reyes"
date: "February 26, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare Environment

Load libraries.

```{r libraries}
library(xlsx)
library(MKmisc)
```


Download data from online Url to subdirectory local to this RStudio project.

```{r download-data}
if (!file.exists("data")) {
  dir.create("data")
}

# !!!This version of file via automated download fails. 
# Remainder of code for assigment assumes manually imported version of .xlsx file located in data subdirectory
# of working (.rproj) directory
am.url <- "http://onlinestatbook.com/2/case_studies/data/angry_moods.xls"
am.file <- "./data/angry_moods.xls"
if (!file.exists(am.file)) {
  download.file(am.url, destfile = am.file)
}
```

Import data.

```{r import-data}
am.df <- read.xlsx(am.file, 
                   sheetName = "angry_moods.txt", 
                   header = TRUE,
                   colClasses = rep("integer", 7))
```

## Q10.

```{r q10}
hist(am.df$Control.Out, main = "Control-Out Scores")
```

## Q11.

```{r q11}
ctlout.mean <- mean(am.df$Control.Out)
ctlout.athete.mean <- mean(am.df$Control.Out[am.df$Sports == 1])
ctlout.nonath.mean <- mean(am.df$Control.Out[am.df$Sports != 1])
```

Mean                         | Result  |
-----------------------------|---------| 
Overall Control-Out Mean     | `r round(ctlout.mean, 2)` |
Athlete Control-Out Mean     | `r round(ctlout.athete.mean, 2)` |
Non-Athlete Control-out Mean | `r round(ctlout.nonath.mean, 2)` |

## Q17.

Box plot of Anger Expression Index by sports participation using Type 6, Lohninger, n+1 calculation of hinges, as
recommended by Introduction to Statistics, David Lane (Accessed at http://onlinestatbook.com/2/index.html).

```{r}
list.q17 <- list(
  Athletes = am.df$Anger_Expression[am.df$Sports == 1], 
  "Non-Athletes" = am.df$Anger_Expression[am.df$Sports != 1])
quantile(am.df$Anger_Expression[am.df$Sports==1], type = 6)
quantile(am.df$Anger_Expression[am.df$Sports!=1], type = 6)
qboxplot(list.q17, type = 6, range = 1.5, main = "Anger Expression Index, Type 6")
```

Box plot of Anger Expression Index by sports participation using Type 7 calculation of hinges, per R default.

```{r q17}
list.q17 <- list(
  Athletes = am.df$Anger_Expression[am.df$Sports == 1], 
  "Non-Athletes" = am.df$Anger_Expression[am.df$Sports != 1])
quantile(am.df$Anger_Expression[am.df$Sports==1], type = 7)
quantile(am.df$Anger_Expression[am.df$Sports!=1], type = 7)
boxplot(list.q17, main = "Anger Expression Index, Type 7")
```

### Analysis

Depending on method of calculation, the lowest Anger Expression Index in the Non-Athletes group is either an outlier 
(Type 7 calculation of hinges) or at least lies near the edge of the lower inner fence.

The non-athletic group showed greater variability than the athletic group, with some extreme lower values (per above) of the
non-athletes showing lower expression of anger than the athletes. However, overall the central tendancy of athletic group
demonstrated a lower Anger Expression Index (interpreted as expressing less anger) than the non-athletic group.

## Q18.

Box plot of Anger Expression Index by gender using Type 6, Lohninger, n+1 calculation of hinges, as
recommended by Introduction to Statistics, David Lane (Accessed at http://onlinestatbook.com/2/index.html).

```{r}
list.q17 <- list(
  Male = am.df$Anger_Expression[am.df$Gender == 1], 
  Female = am.df$Anger_Expression[am.df$Gender == 2])
qboxplot(list.q17, type = 6, main = "Anger Expression Index, Type 6")
```

Box plot of Anger Expression Index by gender using Type 7 calculation of hinges, per R default.

```{r q18}
list.q17 <- list(
  Male = am.df$Anger_Expression[am.df$Gender == 1], 
  Female = am.df$Anger_Expression[am.df$Gender == 2])
boxplot(list.q17, main = "Anger Expression Index, Type 7")
```

## Q20.

Scatter plot to visualize correlation of Control-In and Control-Out.

```{r q20-plot}
plot(am.df$Control.In, am.df$Control.Out)
```

```{r q20-correlation}
x <- am.df$Control.In
y <- am.df$Control.Out
x.bar <- mean(x)
y.bar <- mean(y)

# Calculate Pearson's correlation manually
r.manual <- sum((x - x.bar) * (y - y.bar)) / ((sqrt(sum((x - x.bar)^2))) * (sqrt(sum((y - y.bar)^2))))

# Calculate Pearson's correlation with built-in R function
r <- cor(x, y, method = "pearson")
```

Manual calculation of correlation between Control-In and Control-Out: `r round(r.manual, 4)`

Calculation of correlation between Control-In and Control-Out using cor(): `r round(r, 4)`

```{r q20-significance}
# Calculate t-value using previously calculated correlation, r.
t <- r * (sqrt(nrow(am.df) - 2) / sqrt(1 - r^2))

# Set critical value based on t-table lookup (provided by TA).
critical.value <- 2.374

# Test for significance using t value.
significant <- t > critical.value

# Check built-in cor.test()
cor.test(x, y, alternative = "two.sided", method = "pearson", conf.level = 0.99)

```

Is this correlation statistically significant at the 0.01 level? `r significant`

## Q21.

Since the variable Anger-Out measures the extent that subjects express their anger outwardly, and the variable Control-Out
measures the extent that subjects control the outward expression of anger, I hypothesize that the two variables have a
negative correlation.

Scatter plot to visualize correlation of Control-In and Control-Out.

```{r q21-plot}
plot(am.df$Anger.Out, am.df$Control.Out)
```

```{r q21-correlation}
x <- am.df$Anger.Out
y <- am.df$Control.Out
x.bar <- mean(x)
y.bar <- mean(y)

# Calculate Pearson's correlation manually
r.manual <- sum((x - x.bar) * (y - y.bar)) / ((sqrt(sum((x - x.bar)^2))) * (sqrt(sum((y - y.bar)^2))))

# Calculate Pearson's correlation with built-in R function
r <- cor(x, y, method = "pearson")
```

Manual calculation of correlation between Anger-Out and Control-Out: `r round(r.manual, 4)`

Calculation of correlation between Anger-Out and Control-Out using cor(): `r round(r, 4)`
