---
title: "Introduction to Statistical Concepts, MSCA 31000, Assignment 1b"
author: "Hope Foster-Reyes"
date: "February 26, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare Environment

Load libraries.

```{r libraries}
library(xlsx)
```


Download data from online Url to subdirectory local to this RStudio project.

```{r download-data}
if (!file.exists("data")) {
  dir.create("data")
}

# !!!This version of file fails. Troubleshoot. Below code worked with manually imported version of file.
am.url <- "http://onlinestatbook.com/2/case_studies/data/angry_moods.xls"
am.file <- "./data/angry_moods.xls"
if (!file.exists(am.file)) {
  download.file(am.url, destfile = am.file)
}
```

Import data.

```{r import-data}
am.df <- read.xlsx(am.file, 
                   sheetName = "angry_moods.txt", 
                   header = TRUE,
                   colClasses = rep("integer", 7))
```

## Q10.

```{r q10}
hist(am.df$Control.Out, main = "Control-Out Scores")
```

## Q11.

```{r q11}
ctlout.mean <- mean(am.df$Control.Out)
ctlout.athete.mean <- mean(am.df$Control.Out[am.df$Sports == 1])
ctlout.nonath.mean <- mean(am.df$Control.Out[am.df$Sports != 1])
```

Mean                         | Result  |
-----------------------------|---------| 
Overall Control-Out Mean     | `r round(ctlout.mean, 2)` |
Athlete Control-Out Mean     | `r round(ctlout.athete.mean, 2)` |
Non-Athlete Control-out Mean | `r round(ctlout.nonath.mean, 2)` |

## Q17.

```{r q17}
list.q17 <- list(
  Athletes = am.df$Anger_Expression[am.df$Sports == 1], 
  "Non-Athletes" = am.df$Anger_Expression[am.df$Sports != 1])
boxplot(list.q17, main = "Anger Expression Index")
```

## Q18.
```{r q18}
list.q17 <- list(
  Male = am.df$Anger_Expression[am.df$Gender == 1], 
  Female = am.df$Anger_Expression[am.df$Gender == 2])
boxplot(list.q17, main = "Anger Expression Index")
```

## Q20.

Scatter plot to visualize correlation of Control-In and Control-Out.

```{r q20-plot}
plot(am.df$Control.In, am.df$Control.Out)
```

```{r q20-correlation}
x <- am.df$Control.In
y <- am.df$Control.Out
x.bar <- mean(x)
y.bar <- mean(y)

# Calculate Pearson's correlation manually
r.manual <- sum((x - x.bar) * (y - y.bar)) / ((sqrt(sum((x - x.bar)^2))) * (sqrt(sum((y - y.bar)^2))))

# Calculate Pearson's correlation with built-in R function
r <- cor(x, y, method = "pearson")
```

Manual calculation of correlation between Control-In and Control-Out - `r round(r.manual, 4)`

Calculation of correlation between Control-In and Control-Out using cor() - `r round(r, 4)`

```{r q20-significance}
# Calculate t-value using previously calculated correlation, r.
t <- r * (sqrt(nrow(am.df) - 2) / sqrt(1 - r^2))

# Set critical value based on t-table lookup (provided by TA).
critical.value <- 2.374

# Test for significance using t value.
significant <- t > critical.value

# Check built-in cor.test()
cor.test(x, y, alternative = "two.sided", method = "pearson", conf.level = 0.99)

```

Is this correlation statistically significant at the 0.01 level? `r significant`

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
