---
title: "Introduction to Statistical Concepts, MSCA 31000, Assignment 1b"
author: "Hope Foster-Reyes"
date: "February 26, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare Environment

Load libraries.

```{r}
library(xlsx)
```


Download data from online Url to subdirectory local to this RStudio project.

```{r download-data}
if (!file.exists("data")) {
  dir.create("data")
}

# !!!This version of file fails. Troubleshoot. Below code worked with manually imported version of file.
am.url <- "http://onlinestatbook.com/2/case_studies/data/angry_moods.xls"
am.file <- "./data/angry_moods.xls"
if (!file.exists(am.file)) {
  download.file(am.url, destfile = am.file)
}
```

Import data.

```{r import-data}
am.df <- read.xlsx(am.file, 
                   sheetName = "angry_moods.txt", 
                   header = TRUE,
                   colClasses = rep("integer", 7))
```

## Q10.

```{r q10}
hist(am.df$Control.In)
```

## Q11.

```{r q11}
ctlout.mean <- mean(am.df$Control.Out)
ctlout.athete.mean <- mean(am.df$Control.Out[am.df$Sports == 1])
ctlout.nonath.mean <- mean(am.df$Control.Out[am.df$Sports != 1])
```

Mean                         | Result  |
-----------------------------|---------| 
Overal Control-Out Mean      | `r round(ctlout.mean), 2)` |
Athlete Control-Out Mean     | `r round(ctlout.athete.mean, 2)` |
Non-Athlete Control-out Mean | `r round(ctlout.nonath.mean, 2)` |

## Q17.

list.q17 <- list(
  Athletes = am.df$Anger_Expression[am.df$Sports == 1], 
  "Non-Athletes" = am.df$Anger_Expression[am.df$Sports != 1])
boxplot(list.q17, main = "Anger Expression Index")

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
