---
title: "Introduction to Statistical Concepts, MSCA 31000, Assignment 4"
author: "Hope Foster-Reyes"
date: "March 26, 2016"
output: html_document
header-includes: \usepackage{amsmath}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 14, Q2.

```{r q2}
y.predicted <- function (x, b, a) {
  return((b * x) + a)
}

x.predictor <- function (y, b, a) {
  return((y - a) / b)
}

q2.a <- y.predicted(6, 2, 9)

q2.b <- x.predictor(14, 2, 9)
```

Y' = bX + A

Y' = 2X + 9

### a.

X = 6

Y' = 2X + 9

Y' = (2)6 + 9 = `r q2.a`

### b.

Y' = 14

Y' = 2X + 9

2X = y' - 9

X = (14 - 9)/2 = `r q2.b`

## Chapter 14, Q4.

The standard error of the estimate measures the accuracy of predictions of a linear regression. It is the standard deviation of the errors of prediction.

If:

$x$ = criterion variable

$y$ = predictor variable (actual)

$\hat{y}$ = predicted value

$N$ = size of the population

Then the standard error of the estimate (SEE) for a population can be calculated as follows:

SSE (population) = $\sqrt{\frac{\sum {(y - \hat{y})}^2}{N}}$

## Chapter 14, Q6.

## Chapter 14, Q8.

## Chapter 14, Q10.

## Prepare Environment

Load libraries.

```{r libraries}
library(xlsx)
library(MKmisc)
```

Download data from online Url to subdirectory local to this RStudio project.

```{r download-data}
if (!file.exists("data")) {
  dir.create("data")
}

# !!!This version of file via automated download fails. 
# Remainder of code for assigment assumes manually imported version of .xlsx file located 
# in data subdirectory of working (.rproj) directory
am.url <- "http://onlinestatbook.com/2/case_studies/data/angry_moods.xls"
am.file <- "./data/angry_moods.xls"
if (!file.exists(am.file)) {
  download.file(am.url, destfile = am.file)
}
```

Import data.

```{r import-data}
am.df <- read.xlsx(am.file, 
                   sheetName = "angry_moods.txt", 
                   header = TRUE,
                   colClasses = rep("integer", 7))
```






