---
title: "Introduction to Statistical Concepts, MSCA 31000, Assignment 2b"
author: "Hope Foster-Reyes"
date: "March 3, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare Environment

Load libraries.

```{r libraries}
library(xlsx)
library(MKmisc)
```


Download data from online Url to subdirectory local to this RStudio project.

```{r download-data}
if (!file.exists("data")) {
  dir.create("data")
}

# !!!This version of file via automated download fails. 
# Remainder of code for assigment assumes manually imported version of .xlsx file located in data subdirectory
# of working (.rproj) directory
am.url <- "http://onlinestatbook.com/2/case_studies/data/angry_moods.xls"
am.file <- "./data/angry_moods.xls"
if (!file.exists(am.file)) {
  download.file(am.url, destfile = am.file)
}
```

Import data.

```{r import-data}
am.df <- read.xlsx(am.file, 
                   sheetName = "angry_moods.txt", 
                   header = TRUE,
                   colClasses = rep("integer", 7))
```

## Chapter 7, Q22.

### a.

```{r 7-q22-a}
ae.mean <- mean(am.df$Anger_Expression)
ae.sd <- sd(am.df$Anger_Expression)
```

Anger Expression $\mu$: `r round(ae.mean, 3)`

Anger Expression $\sigma$: `r round(ae.sd, 3)`

### b. & c.

```{r 7-q22-b}
normal.out <- function(p) {
  return(round(qnorm(p, mean = ae.mean, sd = ae.sd), 3))
}

actual.out <- function(p, type = 6) {
  return(round(quantile(am.df$Anger_Expression, p, type = type), 3))
}

```

Percentile  | b. Normal Distribution  | c. Actual (Type 6)   |
------------|-------------------------|----------------------|
25          | `r normal.out(0.25)` | `r actual.out(0.25, 6)` |
50          | `r normal.out(0.50)` | `r actual.out(0.50, 6)` |
75          | `r normal.out(0.75)` | `r actual.out(0.75, 6)` |


## Chapter 9, Q23

```{r 9-q23}
male.count <- sum(am.df$Gender == 1)
female.count <- sum(am.df$Gender == 2)
```

### a. 

Number of men sampled: `r male.count`

### b.

Number of women sampled: `r female.count`

## Chapter 9, Q24

```{r 9-q24}
mf.sample.mean.diff <- mean(am.df$Anger.Out[am.df$Gender == 1]) - mean(am.df$Anger.Out[am.df$Gender == 2])
```

Difference of means between male Anger-Out and female Anger-Out scores in sample: 
$M_{AOm}$ - $M_{AOf}$ = `r mf.sample.mean.diff`

```{r q10}
hist(am.df$Control.Out, main = "Control-Out Scores")
```

## Chapter 9, Q25

```{r 9-q25}
mf.pop.mu.diff <- 2
male.pop.variance <- 20
female.pop.variance <- 20
```

Population difference of means between male Anger-Out and female Anger-Out scores: 

$\mu_{AOm}$ - $\mu_{AOf}$ = `r mf.pop.mu.diff`

Population variance for male Angler-Out scores: $\sigma_{AOm}^2$ = `r male.pop.variance`

Population variance for female Angler-Out scores: $\sigma_{AOf}^2$ = `r female.pop.variance`

### a.

Mean of the sampling distribution of the difference between means:
$\mu_{M_{AOm}}$ - $\mu_{M_{AOf}}$ = $\mu_{AOm}$ - $\mu_{AOf}$ = `r mf.pop.mu.diff`

### b.

```{r}
male.sampling.variance <- male.pop.variance / male.count
female.sampling.variance <- female.pop.variance / female.count
mf.sampling.variance.diff <- male.sampling.variance + female.sampling.variance
```

Standard error of the difference between means:

$\sigma_{M_{AOm} - M_{AOf}}$ = 

### c.



